Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-05-25 awipev-CO2.Rmd
added Omega in time series figure

df <- d_all %>% 
  dplyr::select(datetime, temp_insitu_ctd_1m, temp_insitu_ctd_3m, temp_insitu_ctd_5m, temp_insitu_ctd_7m, temp_insitu_ctd_9m) %>%
  dplyr::mutate(date = format(datetime, '%Y-%m-%d')) %>% 
  group_by(date) %>% 
  summarize(mean1m = mean(temp_insitu_ctd_1m, na.rm = TRUE),
            mean9m = mean(temp_insitu_ctd_9m, na.rm = TRUE),
            range = mean(temp_insitu_ctd_1m, na.rm = TRUE)-mean(temp_insitu_ctd_9m, na.rm = TRUE)) %>% 
  ungroup()
summary(df)
```
There are `r sum(!is.na(df$range))` days with temperature values at both 1 m (0-2 m) and 9 m (> 8 m). The temperature difference between 1 and 9 m ranges between `r min(df$range, na.rm=TRUE)`°C and `r max(df$range, na.rm=TRUE)`°C (mean: `r mean(df$range, na.rm=TRUE)`°C; `r median(df$range, na.rm=TRUE)`°C). The difference with depth is therefore significant.

@@

```{r Plot TA calc TA meas, echo=FALSE, warning=FALSE, message=FALSE,out.width='100%',fig.align='hold'}
TAevents <- as.POSIXct(c("2016-02-26 00:00:00", "2017-02-09 12:00:00", "2017-02-09 12:00:00", "2017-03-10 12:00:00",
              "2018-01-07 00:00:00", "2018-06-20 00:00:00", "2018-07-31 00:00:00", "2018-10-30 18:00:00",
              "2018-10-31 09:00:00", "2019-01-16 09:00:00", "2019-01-26 13:25:00", format(Sys.Date(),
              "%Y-%m-%d %H:%M:%S")))
at_contros_calc_xts <- dplyr::select(d_all, datetime,at_calc, at)
at_contros_calc_xts <-  as.xts(at_contros_calc_xts, order.by =d_all$datetime)
dygraph(at_contros_calc_xts, group = "awipev", main="TA, calculated from sal_mix (blue) and reference (red)", ylab="Total alkalinity") %>%
  #dySeries("AT", label= "Raw TA", color ="blue", strokeWidth = 0, pointSize=2) %>%
  dySeries("at_calc", label="at_calc",color = "blue", strokeWidth = 0, pointSize=2) %>%
  dySeries("at", label="at (ref)",color = "red", strokeWidth = 0, pointSize=4) %>%
  dyEvent(as.POSIXct("2016-05-31 10:00:00", tz="UTC"), "HCl Changed", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2016-06-24 10:00:00", tz="UTC"), "BCG Changed", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2016-02-24 14:00:00", tz="UTC"), "Calibration 1", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2016-09-16 10:00:00", tz="UTC"), "Calibration 2", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2017-03-21 12:00:00", tz="UTC"), "TA analyser changed", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2018-01-08 12:55:00", tz="UTC"), "TA analyser changed #TA-0317-001", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2018-07-31 00:00:00", tz="UTC"), "TA analyser changed #TA-1215-001", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2018-10-30 18:00:00", tz="UTC"), "TA analyser changed #TA-0317-001", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2016-11-15 09:00:00", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2017-02-09 12:20:00", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2017-04-11 06:20:00", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2018-01-22 13:00:12", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2018-05-16 12:39:39", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2018-10-31 10:29:27", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2019-01-26 13:25:20", tz="UTC"), "Calibration mode - CRM", labelLoc = "bottom") %>%
  #dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2"), pointSize = 4) %>%
  dyShading(from = TAevents[1], to = TAevents[2], color = "#EDEDED") %>%
  dyShading(from = TAevents[3], to = TAevents[4], color = "#C2C2C2") %>%
  dyShading(from = TAevents[5], to = TAevents[6], color = "#EDEDED") %>%
  dyShading(from = TAevents[7], to = TAevents[8], color = "#C2C2C2") %>%
  dyShading(from = TAevents[9], to = TAevents[10], color = "#EDEDED") %>%
  dyShading(from = TAevents[11], to = TAevents[12], color = "#C2C2C2") %>%
  #dyAxis("y",valueRange = c(2000, 2650)) %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions( strokeWidth= 0) %>%
  dyLegend(show = "follow")   %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE,useDataTimezone = TRUE) %>%
    dyRangeSelector(height = 30, dateWindow= NULL)

@@

pdelta_at_snapo
pdelta_at_snapo_dygraph
p_at_ts
pdelta_ct_snapo
pdelta_ct_snapo_dygraph
p_ct_ts
```
```{r save all, echo=FALSE, message= FALSE, warning=FALSE}
# d_hour is the real data set that is read at the beginning of the script
# d_all comprises additional columns calculated in this script
saveRDS(file= paste0(path, "fb_data/d_all.rds"), d_all)
```

@@

xlim <- c(as.POSIXct("2015-07-01 00:00:00"), as.POSIXct("2020-07-01 00:00:00"))
# select time series data
d <- as_tibble(d_all) %>%
  dplyr::select(datetime, temp_insitu_11m, sal_mix, pco2_corr, pHint_tot_sf, ph_dur_corr, at_calc)
# time series temp
ts_temp <- d %>%
  dplyr::filter(!is.na(temp_insitu_11m)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = temp_insitu_11m), col = "blue", size = 0.15) +
  labs(title = "", x = "", y = "Temp. 11 m (°C)") +
  Mytheme(size_labs = 8)
print(ts_temp) 
# time series sal
ts_sal <- d %>%
  dplyr::filter(!is.na(sal_mix)) %>% 
ggplot() +
  geom_point(aes(x = datetime, y = sal_mix), col = "blue", size = 0.15) + 
  labs(title = "", x = "", y = "Salinity") +
  Mytheme(size_labs = 8)
print(ts_sal) 
# time series pco2
ts_pco2 <- d %>%
  dplyr::filter(!is.na(pco2_corr)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = pco2_corr), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "", y = expression(paste("pC", O[2], " (", mu, "atm)"))) +
  Mytheme(size_labs = 8)
print(ts_pco2) 
# time series pH
ts_ph <- d %>%
  dplyr::select(datetime, pHint_tot_sf, ph_dur_corr) %>%
  dplyr::filter(!is.na(pHint_tot_sf)) %>%
  pivot_longer(-datetime, names_to = "pH", values_to = "value") %>%
  ggplot() +
  geom_point(aes(x = datetime, y = value, colour = pH), size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "",
       x = "",
       y = expression(paste("p", H[T]))) +
  # scale_color_manual(
  #   values = c("blue", "red"),
  #   name = "",
  #   breaks = c("pHint_tot_sf", "ph_dur_corr"),
  #   labels = c("SeaFET", "Durafet")
  # ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.35,
    label = "seaFET",
    color = "blue", size = 2
  ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.25,
    label = "Durafet",
    color = "red", size = 2
  ) +
  Mytheme(size_labs = 8) +
  theme(legend.position = "none")
print(ts_ph) 
# time series AT
ts_at <- d %>%
  dplyr::filter(!is.na(at_calc)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = at_calc), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = expression(paste(italic(A)[T]," ","(",mu, mol.kg^-1,")"))) +
  Mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_at)

#assemble time-series figure
cp <- cowplot::plot_grid(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, 
                   align = "v",
                   #labels = "AUTO",
                   ncol = 1
                   #width = 18,
                   #units = "cm"
                   )
print(cp)
cowplot::ggsave2(filename = "figures/ts_gg.png", plot = cp, height = 20, units = "cm")

>>>>

df <- d_all %>% 
  dplyr::select(datetime, temp_insitu_ctd_1m, temp_insitu_ctd_3m, temp_insitu_ctd_5m, temp_insitu_ctd_7m, temp_insitu_ctd_9m) %>%
  dplyr::mutate(date = format(datetime, '%Y-%m-%d')) %>% 
  group_by(date) %>% 
  summarize(mean1m = mean(temp_insitu_ctd_1m, na.rm = TRUE),
            mean9m = mean(temp_insitu_ctd_9m, na.rm = TRUE),
            range = mean(temp_insitu_ctd_1m, na.rm = TRUE)-mean(temp_insitu_ctd_9m, na.rm = TRUE)) %>% 
  ungroup()
```
There are `r sum(!is.na(df$range))` days with temperature values at both 1 m (0-2 m) and 9 m (> 8 m). The temperature difference between 1 and 9 m ranges between `r min(df$range, na.rm=TRUE)`°C and `r max(df$range, na.rm=TRUE)`°C (mean: `r mean(df$range, na.rm=TRUE)`°C; `r median(df$range, na.rm=TRUE)`°C). The difference with depth is therefore significant.

@@

```{r Plot TA calc TA meas, echo=FALSE, warning=FALSE, message=FALSE,out.width='100%',fig.align='hold'}
TAevents <- as.POSIXct(c("2016-02-26 00:00:00", "2017-02-09 12:00:00", "2017-02-09 12:00:00", "2017-03-10 12:00:00",
              "2018-01-07 00:00:00", "2018-06-20 00:00:00", "2018-07-31 00:00:00", "2018-10-30 18:00:00",
              "2018-10-31 09:00:00", "2019-01-16 09:00:00", "2019-01-26 13:25:00", format(Sys.Date(),
              "%Y-%m-%d %H:%M:%S")))
at_contros_calc_xts <- dplyr::select(d_all, datetime, at_calc, at)
at_contros_calc_xts <-  as.xts(at_contros_calc_xts, order.by =d_all$datetime)
dygraph(at_contros_calc_xts, group = "awipev", main="TA, calculated from sal_mix (blue) and reference (red)", ylab="Total alkalinity") %>%
  #dySeries("AT", label= "Raw TA", color ="blue", strokeWidth = 0, pointSize=2) %>%
  dySeries("at_calc", label="at_calc",color = "blue", strokeWidth = 0, pointSize=2) %>%
  dySeries("at", label="at (ref)",color = "red", strokeWidth = 0, pointSize=4) %>%
  dyEvent(as.POSIXct("2016-05-31 10:00:00", tz="UTC"), "HCl Changed", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2016-06-24 10:00:00", tz="UTC"), "BCG Changed", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2016-02-24 14:00:00", tz="UTC"), "Calibration 1", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2016-09-16 10:00:00", tz="UTC"), "Calibration 2", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2017-03-21 12:00:00", tz="UTC"), "TA analyser changed", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2018-01-08 12:55:00", tz="UTC"), "TA analyser changed #TA-0317-001", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2018-07-31 00:00:00", tz="UTC"), "TA analyser changed #TA-1215-001", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2018-10-30 18:00:00", tz="UTC"), "TA analyser changed #TA-0317-001", labelLoc = "top", color="red") %>%
  dyEvent(as.POSIXct("2016-11-15 09:00:00", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2017-02-09 12:20:00", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2017-04-11 06:20:00", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2018-01-22 13:00:12", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2018-05-16 12:39:39", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2018-10-31 10:29:27", tz="UTC"), "CRM measurement", labelLoc = "bottom") %>%
  dyEvent(as.POSIXct("2019-01-26 13:25:20", tz="UTC"), "Calibration mode - CRM", labelLoc = "bottom") %>%
  #dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2"), pointSize = 4) %>%
  dyShading(from = TAevents[1], to = TAevents[2], color = "#EDEDED") %>%
  dyShading(from = TAevents[3], to = TAevents[4], color = "#C2C2C2") %>%
  dyShading(from = TAevents[5], to = TAevents[6], color = "#EDEDED") %>%
  dyShading(from = TAevents[7], to = TAevents[8], color = "#C2C2C2") %>%
  dyShading(from = TAevents[9], to = TAevents[10], color = "#EDEDED") %>%
  dyShading(from = TAevents[11], to = TAevents[12], color = "#C2C2C2") %>%
  #dyAxis("y",valueRange = c(2000, 2650)) %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions( strokeWidth= 0) %>%
  dyLegend(show = "follow")   %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE,useDataTimezone = TRUE) %>%
    dyRangeSelector(height = 30, dateWindow= NULL)

@@

pdelta_at_snapo
pdelta_at_snapo_dygraph
p_at_ts
pdelta_ct_snapo
pdelta_ct_snapo_dygraph
p_ct_ts
```

```{r save all, echo=FALSE, message= FALSE, warning=FALSE}
# d_hour is the real data set that is read at the beginning of the script
# d_all comprises additional columns calculated in this script
d_all$at <- ifelse(is.nan(d_all$at), NA, d_all$at) # replace all NaN by NA
d_all$ct <- ifelse(is.nan(d_all$at), NA, d_all$ct) # replace all NaN by NA
saveRDS(file= paste0(path, "fb_data/d_all.rds"), d_all)
```

@@

xlim <- c(as.POSIXct("2015-07-01 00:00:00"), as.POSIXct("2020-07-01 00:00:00"))
# calculate omeaga arag
z <- d_all %>% 
  dplyr::filter(!(is.na(pco2_corr) | is.na(at_calc) | is.na(sal_mix) |is.na(temp_insitu_11m) | is.na(pressure_insitu_ctd)))
oa_tbf <- as_tibble(
  carb(24, z$pco2_corr, z$at_calc*1e-6, S = z$sal_mix, T = z$temp_insitu_11m, 
                          P = z$pressure_insitu_ctd/10, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", 
                          ks = "d", pHscale = "T", b = "u74")
) %>% 
  dplyr::mutate(datetime = z$datetime,
                oa = OmegaAragonite) %>% 
dplyr::select(datetime, oa)
# select time series data
d <- as_tibble(d_all) %>%
  dplyr::select(datetime, pressure_insitu_ctd, temp_insitu_11m, sal_mix, pco2_corr, pHint_tot_sf, 
                ph_dur_corr, at_calc)
d <- left_join(d, oa_tbf, by = 'datetime')

# time series temp
ts_temp <- d %>%
  dplyr::filter(!is.na(temp_insitu_11m)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = temp_insitu_11m), col = "blue", size = 0.15) +
  labs(title = "", x = "", y = "Temp. 11 m (°C)") +
  Mytheme(size_labs = 8)
print(ts_temp) 
# time series sal
ts_sal <- d %>%
  dplyr::filter(!is.na(sal_mix)) %>% 
ggplot() +
  geom_point(aes(x = datetime, y = sal_mix), col = "blue", size = 0.15) + 
  labs(title = "", x = "", y = "Salinity") +
  Mytheme(size_labs = 8)
print(ts_sal) 
# time series pco2
ts_pco2 <- d %>%
  dplyr::filter(!is.na(pco2_corr)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = pco2_corr), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "", y = expression(paste("pC", O[2], " (", mu, "atm)"))) +
  Mytheme(size_labs = 8)
print(ts_pco2) 
# time series pH
ts_ph <- d %>%
  dplyr::select(datetime, pHint_tot_sf, ph_dur_corr) %>%
  dplyr::filter(!is.na(pHint_tot_sf)) %>%
  pivot_longer(-datetime, names_to = "pH", values_to = "value") %>%
  ggplot() +
  geom_point(aes(x = datetime, y = value, colour = pH), size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "",
       x = "",
       y = expression(paste("p", H[T]))) +
  # scale_color_manual(
  #   values = c("blue", "red"),
  #   name = "",
  #   breaks = c("pHint_tot_sf", "ph_dur_corr"),
  #   labels = c("SeaFET", "Durafet")
  # ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.35,
    label = "seaFET",
    color = "blue", size = 2
  ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.25,
    label = "Durafet",
    color = "red", size = 2
  ) +
  Mytheme(size_labs = 8) +
  theme(legend.position = "none")
print(ts_ph) 
# time series AT
ts_at <- d %>%
  dplyr::filter(!is.na(at_calc)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = at_calc), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = expression(paste(italic(A)[T]," ","(",mu, mol.kg^-1,")"))) +
  Mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_at) 
# time series Omega
ts_oa <- d %>%
  dplyr::filter(!is.na(oa)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = oa), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = "Omega aragonite") +
  Mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_oa) 
#assemble time-series figure
cp <- cowplot::plot_grid(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, ts_oa,
                   align = "v",
                   #labels = "AUTO",
                   ncol = 1
                   #width = 18,
                   #units = "cm"
                   )
print(cp)
cowplot::ggsave2(filename = "figures/ts_gg.png", plot = cp, height = 20, units = "cm")
